
---
title: "UMAP projection of interfollicular epidermal cells in unwounded (UW) sample"
author: "Suoqin Jin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mainfont: Arial
vignette: >
  %\VignetteIndexEntry{Pseudotime analysis of interfollicular epidermal cells in UW sample}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

  ```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  root.dir = './'
)
```

This walkthrough performs UMAP projection of interfollicular epidermal cells in unwounded (UW) sample, including proliferative and non-proliferative basal cells and spinous cells

The reduced UMAP space will be saved and then used for pseudotime analysis in scEpath (https://github.com/sqjin/scEpath)


## Load the required libraries
```{r message=FALSE,warning=FALSE}
library("Seurat", lib.loc="/Users/suoqinjin/Documents/Rlibrary") # Seurat version 2.3.4
library(sva) # sva version 3.30.1
```

## Load the raw count data and cell group information

```{r}
setwd("/Users/suoqinjin/Google Drive/projects/project4_Dai/revisionR1/codes_submit")
load("data_pseudotime.RData")
```

## Setup the Seurat Object
```{r}
# Initialize the Seurat object with the raw (non-normalized data).
w10x_UW <- CreateSeuratObject(raw.data = data_pseudotime$raw.data, meta.data = data_pseudotime$meta.data)
# set cell identity
w10x_UW <- SetAllIdent(w10x_UW, id = 'clusters')
```

## Normalizing the data
```{r, results='asis'}
w10x_UW <- NormalizeData(object = w10x_UW, normalization.method = "LogNormalize", scale.factor = 10000)
```

## Perform batch correction
```{r}
# batch-correction using ComBat
batchid = rep(1,length(w10x_UW@meta.data$batch))
batchid[w10x_UW@meta.data$batch=="bs_2"] = 2
names(batchid) =colnames(w10x_UW@data)
m = as.data.frame(as.matrix(w10x_UW@data))
m = m[rowSums(m)>0,]
com = ComBat(as.matrix(m), batchid, prior.plots=FALSE, par.prior=TRUE)
w10x_UW@data = Matrix(as.matrix(com))
```

## Find highly variable genes and scale the data
```{r}
w10x_UW <- FindVariableGenes(object = w10x_UW, mean.function = ExpMean, dispersion.function = LogVMR, x.low.cutoff = 0.01, x.high.cutoff = 4, y.cutoff = 0.6, do.plot = FALSE)
w10x_UW <- ScaleData(w10x_UW)
```

## Run PCA
```{r, fig.width=5,fig.height = 3, fig.wide = TRUE, fig.align = "center"}
w10x_UW <- RunPCA(w10x_UW, pc.genes = w10x_UW@var.genes, pcs.compute = 40, do.print = FALSE)
PCElbowPlot(w10x_UW, num.pc = 40)
```

## Run UMAP
```{r}
# use top six PCs with significant deviations
w10x_UW <- RunUMAP(w10x_UW, reduction.use = "pca", dims.use = 1:6,min_dist = 0.35)
```

## Visualize cells in the reduced UMAP space (colored by clusters)
```{r, fig.width=5,fig.height = 3.5,  fig.wide = TRUE, fig.align = "center"}
color.use <- c('#ff1622','#00b938','#3657df','#603A91','#F29233','#D6B239')
DimPlot(w10x_UW, reduction.use = "umap", cols.use = color.use, pt.size = 0.2)
```

## Visualize cells in the reduced UMAP space (colored by sample id)
```{r, fig.width=4.5,fig.height = 3.5,  fig.wide = TRUE, fig.align = "center"}
DimPlot(w10x_UW, reduction.use = "umap",pt.size = 0.2, group.by = "batch", cols.use = c("#F8766D","#C49A00"))
```

## Save data for further pseudotime analysis in scEpath and Monocle 3
```{r}
# save data for pseudotime analysis in scEpath
write.table(w10x_UW@dr$umap@cell.embeddings,file = "projectedData_ump_UW_IFE.txt",sep = '\t')
write.table(w10x_UW@var.genes, file = "HVG_UW_IFE.txt", sep ="\t", row.names = F, col.names = F)
write.table(as.matrix(w10x_UW@raw.data),file = "rawData_UW_IFE.txt",sep = '\t')
write.table(w10x_UW@ident, file = "clusters_UW_IFE.txt",sep = '\t')
# save the workspace for pseudotime analysis in Monocle 3
save(w10x_UW, file = "seurat_pseudotimeAnalysis_UW_IFE.RData")

```


